---
title: "Analysis"
author:
- Veronica Price-Jones
- Lien Reyserhove
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
subtitle: ''
---

(short description of the goal of this document)

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(readxl)         # To read Excel files
library(janitor)        # To clean input data
library(ordinal)        # To perform clm
```

# Read source data

Create a data frame `input_data` from the source data:
 
```{r}
input_data <- read_csv(file = here::here("data", "processed_data.csv")) 
```

# Clean data

Remove empty rows (if present) and clean column headers:

```{r}
input_data %<>%
  remove_empty("rows") %>%       # Remove empty rows
  clean_names()                  # Have sensible (lowercase) column names
```

Show cleaned column names:

```{r}
colnames(input_data)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Pre-processing

## Number of invasive alien records

Rename variable `number_of_alien_or_invasive_alien_species_records` to `no_ias_records` to simplify the analysis:

```{r}
input_data <- input_data %>% 
  rename(no_ias_records = number_of_alien_or_invasive_alien_species_records) 
```

Inspect different values of `no_ias_records`:

```{r}
unique(input_data$no_ias_records)
```

Change `Info not available` and `Not applicable` to NA:

```{r}
input_data$no_ias_records <- na_if(input_data$no_ias_records, "Info not available") 
input_data$no_ias_records <- na_if(input_data$no_ias_records, "Not applicable") 
input_data$no_ias_records <- na_if(input_data$no_ias_records, "NA") 
```

Define `no_ias_records` as an ordered factor:

```{r}
input_data$no_ias_records <- factor(input_data $ no_ias_records, ordered = TRUE)
```

Re-order `no_ias_records`:

```{r}
input_data$no_ias_records <- 
  ordered(input_data$no_ias_records, 
    levels = c("1 - 50", "51 - 100",
                "101 - 500", "501 - 1000", 
                "1001 - 5000", "5001 - 10000",
                "10001 - 50000", "50001 - 100000", 
                "100001 - 500000","500001 - 1000000",
                "> 1000001"))
```

Verify order:

```{r}
levels(input_data$no_ias_records)
```


# Analysis

## Exploratory analysis of selected project parameters

Put here all relevant code related to the analyses

## Geographic difference in project scope

Put here all relevant code related to the analyses

## Project performance

project design
use of social media
level of skill required
expected contribution frequencey
provision of guidelines
... training
... sighting maps
... active informing
... feedback
... support


number of invasive alien records
number of publications 
number of attendees

## Number of invasive alien records
### Cumulative Link Model 

Inspect dependent variables on percentage of `NA`:

```{r}
input_data %>% 
  select(design, duration, sm_yn, skills, contribution_frequency, guidelines, training_yn, design, map, active_informing, feedback, support) %>% 
  map(~mean(is.na(.)))
```

Conclusion: variables `feedback` (22%) and `map` (11%) have the highest percentages of missing values


Cumulative link model:

```{r}
clm_records <- 
  clm(
    no_ias_records ~ 
      duration + design + sm_yn + 
      skills + contribution_frequency + guidelines + 
      training_yn + map + active_informing + feedback + support,
    data = input_data)
```

Model output:

```{r}
summary(clm_records)
```

Output with p-values:

```{r}
anova(clm_records)
```

Conclusion: 
- significant positive effect of map; significant negative effect of feedback and of low & no skill requirement
- 64 deleted observations due to empty values in the dependent and independent variables


## No IAS records


  
  
## Number of participants

Inspect different values of `number_of_participants`:

```{r}
unique(input_data$number_of_participants)
```

Change `Info not available` and `Not applicable` to NA:

```{r}
input_data$number_of_participants <- na_if(input_data$number_of_participants, "Info not available") 
input_data$number_of_participants <- na_if(input_data$number_of_participants, "Not applicable") 
input_data$number_of_participants <- na_if(input_data$number_of_participants, "NA") 
```

Define `number_of_participants` as an ordered factor:

```{r}
input_data$number_of_participants <- factor(input_data $ number_of_participants, ordered = TRUE)
```

Re-order `number_of_participants`:

```{r}
input_data$number_of_participants <- ordered (input_data$number_of_participants, levels = c("1 - 50", "51 - 100", 
"101 - 500", "501 - 1000", "1001 - 2000", "2001 - 5000", "5001 - 10000", "> 10001"))
```

Verify order:

```{r}
levels(input_data$number_of_participants)
```

### Cumulative Link Model

The following variables are specified for analysis:
- `design`
- `duration`
- `sm_yn`
- `skills`
- `contribution_frequency`
- `guidelines`
- `training_yn`
- `map`
- `active_informing`
- `feedback`
- `support`

Cumulative link model:

```{r}
clm_participants <- 
  clm(
    number_of_participants ~ duration + design + sm_yn + species_id + guidelines + training_yn + map + active_informing + feedback,
    data = input_data)
```

Model output:

```{r}
summary(clm_participants)
```  

```{r}
anova(clm_participants)
```

Conclusion: 
- significant positive effect of map
- 61 deleted observations



  
## Number of publications

Inspect different values of `number_of_publications`:

```{r}
unique(input_data$number_of_publications)
```

Change `Unknown` and `Not applicable` to NA:

```{r}
input_data$number_of_publications <- na_if(input_data$number_of_publications, "Unknown") 
input_data$number_of_publications <- na_if(input_data$number_of_publications, "Not applicable") 
input_data$number_of_publications <- na_if(input_data$number_of_publications, "NA") 
```

Define `number_of_publications` as an ordered factor:

```{r}
input_data$number_of_publications <- factor(input_data $ number_of_publications, ordered = TRUE)
```

Re-order `number_of_participants`:

```{r}
input_data$number_of_publications <- ordered (input_data$number_of_publications, levels = c("0", "1 to 5", "6 to 10", "11 to 20", "21 - 50", "> 100"))
```

Verify order:

```{r}
str(input_data$number_of_publications)
```

### Cumulative Link Model

The following variables are specified for analysis:
- `design`
- `duration`
- `sm_yn`
- `skills`
- `contribution_frequency`
- `guidelines`
- `training_yn`
- `map`
- `active_informing`
- `feedback`
- `support`

Cumulative link model:

```{r}
clm_publications <- 
  clm(
    number_of_publications ~ duration + design + sm_yn + species_id + guidelines + training_yn + map + active_informing + feedback,
    data = input_data)
```

Model output:

```{r}
summary(clm_publications)
```  

```{r}
anova(publications)
```

Conclusion:
- significant positive effect of training and map; significant negative effect of support and limited or no skill requirement
- 63 deleted observations
