---
title: "Analysis"
author:
- Veronica Price-Jones
- Lien Reyserhove
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
subtitle: ''
---

(short description of the goal of this document)

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(readxl)         # To read Excel files
library(janitor)        # To clean input data
library(ordinal)        # To perform clm
```

# Read source data

Create a data frame `input_data` from the source data:
 
```{r}
input_data <- read_csv(file = here::here("data", "processed_data.csv")) 
```

# Clean data

Remove empty rows (if present) and clean column headers:

```{r}
input_data %<>%
  remove_empty("rows") %>%       # Remove empty rows
  clean_names()                  # Have sensible (lowercase) column names
```

Show cleaned column names:

```{r}
colnames(input_data)
```

Preview data:

```{r}
input_data %>% head(n = 5)
```

# Pre-processing

## Number of invasive alien records

Rename variable `number_of_alien_or_invasive_alien_species_records` to `no_ias_records` to simplify the analysis:

```{r}
input_data <- input_data %>% 
  rename(no_ias_records = number_of_alien_or_invasive_alien_species_records) 
```

Inspect different values of `no_ias_records`:

```{r}
unique(input_data$no_ias_records)
```

Change `Info not available` and `Not applicable` to NA:

```{r}
input_data$no_ias_records <- na_if(input_data$no_ias_records, "Info not available") 
input_data$no_ias_records <- na_if(input_data$no_ias_records, "Not applicable") 
input_data$no_ias_records <- na_if(input_data$no_ias_records, "NA") 
```

Define `no_ias_records` as an ordered factor:

```{r}
input_data$no_ias_records <- factor(input_data $ no_ias_records, ordered = TRUE)
```

Re-order `no_ias_records`:

```{r}
input_data$no_ias_records <- 
  ordered(input_data$no_ias_records, 
    levels = c("1 - 50", "51 - 100",
                "101 - 500", "501 - 1000", 
                "1001 - 5000", "5001 - 10000",
                "10001 - 50000", "50001 - 100000", 
                "100001 - 500000","500001 - 1000000",
                "> 1000001"))
```

Verify order:

```{r}
levels(input_data$no_ias_records)
```

# Analysds

## Exploratory analysis of selected project parameters

We first test for association between target audience and environment

Environment x Audience
Regions
Aim
Data type
Taxon







## Geographic difference in project scope

Put here all relevant code related to the analyses

## Project performance

### Inspect independent variables

The following independent variables were selected for the analyses:

- duration (`duration`)
- project design (`design`)
- use of social media (`sm_yn`)
- level of skill required (`skills`)
- expected contribution frequencey (`contribution_frequency`)
- provision of guidelines (`guidelines`)
- training (`training_yn`)
- sighting maps (`map`)
- active informing (`active_informing`)
- feedback (`feedback`)
- support (`support`) 


Inspect dependent variables on percentage of `NA`:

```{r}
input_data %>% 
  select(duration, design, sm_yn,  skills, contribution_frequency, guidelines, training_yn, map, active_informing, feedback, support) %>% 
  map(~mean(is.na(.)))
```

Conclusion: variables `feedback` (22%) and `map` (11%) have the highest percentages of missing values


### Number of invasive alien records 

Cumulative link model:

```{r}
clm_ias_records <- 
  clm(
    no_ias_records ~ 
      duration + design + sm_yn + 
      skills + contribution_frequency + guidelines + 
      training_yn + map + active_informing + feedback + 
      support,
    data = input_data)
```

Model output:

```{r}
summary(clm_records)
```
Output with p-values:

```{r}
anova(clm_records)
```


## No IAS records

### Secondary Analysis: Engagement

The following variables are specified for analysis:
- `duration`
- `eng_website`
- `eng_sm`
- `eng_live`
- `eng_newsletter`
- `eng_school`
- `eng_exhibitions`
- `eng_blitz`
- `eng_gaming`
- `eng_other`

Convert predictors to factors:

```{r}
input_data$eng_website <- factor(input_data $ eng_website)
input_data$eng_sm <- factor(input_data $ eng_sm)
input_data$eng_live <- factor(input_data $ eng_live)
input_data$eng_newsletter <- factor(input_data $ eng_newsletter)
input_data$eng_school <- factor(input_data $ eng_school)
input_data$eng_exhibitions <- factor(input_data $ eng_exhibitions)
input_data$eng_blitz <- factor(input_data $ eng_blitz)
input_data$eng_gaming <- factor(input_data $ eng_gaming)
input_data$eng_other <- factor(input_data $ eng_other)
```

Cumulative link model:

```{r}
clm_eng_records <- 
  clm(
    no_ias_records ~ duration + eng_website + eng_sm + eng_live + eng_newsletter + eng_school + eng_exhibitions + eng_blitz + eng_gaming + eng_other,
    data = input_data)
```

Model output:

```{r}
summary(clm_eng_records)
```

```{r}
anova(clm_eng_records)
```

Conclusion: 
- no significant results
- 50 deleted observations


### Secondary Analysis: Social media
The following variables are specified for analysis:
- `duration`
- `sm_facebook`
- `sm_twitter`
- `sm_instagram`
- `sm_you_tube`

Convert predictors to factors:

```{r}
input_data$sm_facebook <- factor(input_data $ sm_facebook)
input_data$sm_instagram <- factor(input_data $ sm_instagram)
input_data$sm_twitter <- factor(input_data $ sm_twitter)
input_data$sm_you_tube <- factor(input_data $ sm_you_tube)
input_data$sm_other <- factor(input_data $ sm_other)
```

Cumulative link model:

```{r}
clm_sm_records <- 
  clm(
    no_ias_records ~ duration + sm_facebook + sm_twitter + sm_instagram + sm_you_tube,
    data = input_data)
```

Model output:

```{r}
summary(clm_sm_records)

```{r}
anova(clm_sm_records)
```

Conclusion: 
- no significant results
- 54 deleted observations

  
  
## Number of participants

Inspect different values of `number_of_participants`:

```{r}
unique(input_data$number_of_participants)
```

Change `Info not available` and `Not applicable` to NA:

```{r}
input_data$number_of_participants <- na_if(input_data$number_of_participants, "Info not available") 
input_data$number_of_participants <- na_if(input_data$number_of_participants, "Not applicable") 
input_data$number_of_participants <- na_if(input_data$number_of_participants, "NA") 
```

Define `number_of_participants` as an ordered factor:

```{r}
input_data$number_of_participants <- factor(input_data $ number_of_participants, ordered = TRUE)
```

Re-order `number_of_participants`:

```{r}
input_data$number_of_participants <- ordered (input_data$number_of_participants, levels = c("1 - 50", "51 - 100", 
"101 - 500", "501 - 1000", "1001 - 2000", "2001 - 5000", "5001 - 10000", "> 10001"))
```

Verify order:

```{r}
levels(input_data$number_of_participants)
```

### Primary Analysis

The following variables are specified for analysis:
- `design`
- `duration`
- `sm_yn`
- `skills`
- `contribution_frequency`
- `guidelines`
- `training_yn`
- `map`
- `active_informing`
- `feedback`
- `support`

Cumulative link model:

```{r}
clm_participants <- 
  clm(
    number_of_participants ~ duration + design + sm_yn + species_id + guidelines + training_yn + map + active_informing + feedback,
    data = input_data)
```

Model output:

```{r}
summary(clm_participants)
```  

```{r}
anova(clm_participants)
```

Conclusion: 
- significant positive effect of map
- 61 deleted observations

### Secondary Analysis: Engagement

The following variables are specified for analysis:
- `duration`
- `eng_website`
- `eng_sm`
- `eng_live`
- `eng_newsletter`
- `eng_school`
- `eng_exhibitions`
- `eng_blitz`
- `eng_gaming`
- `eng_other`

Cumulative link model:

```{r}
clm_eng_participants <- 
  clm(
    number_of_participants ~ duration + eng_website + eng_sm + eng_live + eng_newsletter + eng_school + eng_exhibitions + eng_blitz + eng_gaming + eng_other,
    data = input_data)
```

Model output:

```{r}
summary(clm_eng_participants)
```  

```{r}
anova(clm_eng_participants)
```


### Secondary Analysis: Social Media

The following variables are specified for analysis:
- `duration`
- `sm_facebook`
- `sm_twitter`
- `sm_instagram`
- `sm_you_tube`

Cumulative link model:

```{r}
clm_sm_participants <- 
  clm(
    number_of_participants ~ duration + sm_facebook + sm_twitter + sm_instagram + sm_you_tube,
    data = input_data)
```

Model output:

```{r}
summary(clm_sm_participants)
```  

```{r}
anova(clm_sm_participants)
```

Conclusion:
- no significant results
- 46 deleted observations


  
## Number of publications

Inspect different values of `number_of_publications`:

```{r}
unique(input_data$number_of_publications)
```

Change `Unknown` and `Not applicable` to NA:

```{r}
input_data$number_of_publications <- na_if(input_data$number_of_publications, "Unknown") 
input_data$number_of_publications <- na_if(input_data$number_of_publications, "Not applicable") 
input_data$number_of_publications <- na_if(input_data$number_of_publications, "NA") 
```

Define `number_of_publications` as an ordered factor:

```{r}
input_data$number_of_publications <- factor(input_data $ number_of_publications, ordered = TRUE)
```

Re-order `number_of_participants`:

```{r}
input_data$number_of_publications <- ordered (input_data$number_of_publications, levels = c("0", "1 to 5", "6 to 10", "11 to 20", "21 - 50", "> 100"))
```

Verify order:

```{r}
str(input_data$number_of_publications)
```

### Primary Analysis

The following variables are specified for analysis:
- `design`
- `duration`
- `sm_yn`
- `skills`
- `contribution_frequency`
- `guidelines`
- `training_yn`
- `map`
- `active_informing`
- `feedback`
- `support`

Cumulative link model:

```{r}
clm_publications <- 
  clm(
    number_of_publications ~ duration + design + sm_yn + species_id + guidelines + training_yn + map + active_informing + feedback,
    data = input_data)
```

Model output:

```{r}
summary(clm_publications)
```  

```{r}
anova(publications)
```

Conclusion:
- significant positive effect of training and map; significant negative effect of support and limited or no skill requirement
- 63 deleted observations

### Secondary Analysis: Engagement

The following variables are specified for analysis:
- `duration`
- `eng_website`
- `eng_sm`
- `eng_live`
- `eng_newsletter`
- `eng_school`
- `eng_exhibitions`
- `eng_blitz`
- `eng_gaming`
- `eng_other`


Cumulative link model:

```{r}
clm_eng_publications <- 
  clm(
    number_of_publications ~ duration + eng_website + eng_sm + eng_live + eng_newsletter + eng_school + eng_exhibitions + eng_blitz + eng_gaming + eng_other,
    data = input_data)
```

Model output:

```{r}
summary(clm_eng_publications)
```  

```{r}
anova(clm_eng_publications)
```

Conclusion:
- significant positive effect of exhibition
- 48 deleted observations


### Secondary Analysis: Social Media

The following variables are specified for analysis:
- `duration`
- `sm_facebook`
- `sm_twitter`
- `sm_instagram`
- `sm_you_tube`

Cumulative link model:

```{r}
clm_sm_publications <- 
  clm(
    number_of_publications ~ duration + sm_facebook + sm_twitter + sm_instagram + sm_you_tube,
    data = input_data)
```

Model output:

```{r}
summary(clm_sm_publications)
```  

```{r}
anova(clm_sm_publications)
```

Conclusion:
- no significant results
- 52 deleted observations  
