---
title: "Code to reproduce barplots"
author:
- Quentin Groom  
- Veronica Price-Jones
- Lien Reyserhove
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
subtitle: ''
---

# General settings

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Load libraries:

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(readxl)         # To read Excel files
library(dplyr)          # To clean input data
library(readr)
library(janitor)
```

# Import files

```{r}
appsProjects <- 
  read_delim(
    file = here::here("data", "appsProjects.txt"),
    escape_double = FALSE, 
    trim_ws = TRUE)
                           
                           
newProjectsThatMapAlienSpecies <- 
  read_delim(
    file = here::here("data", "newProjectsThatMapAlienSpecies.txt"), 
    escape_double = FALSE, 
    trim_ws = TRUE)

newProjectsWithPolicyRelevance <- 
  read_delim(
    file = here::here("data", "newProjectsWithPolicyRelevance.txt"), 
    escape_double = FALSE, 
    trim_ws = TRUE)

numberOfNewProjects <- 
  read_delim(
    file = here::here("data","numberOfNewProjects.txt"), 
    escape_double = FALSE, 
    trim_ws = TRUE)

numberOfParticiProjects <- 
  read_delim(
    file = here::here("data","numberOfParticiProjects.txt"),
    escape_double = FALSE, 
    trim_ws = TRUE)

numberOfPubsProjects <- 
  read_delim(
    file = here::here("data","numberOfPubsProjects.txt"), 
    escape_double = FALSE, 
    trim_ws = TRUE)

numberOfRecsProjects <- 
  read_delim(
    here::here("data","numberOfRecsProjects.txt"), 
    escape_double = FALSE, 
    trim_ws = TRUE)
```

Clean names:

```{r}
appsProjects <- appsProjects %>% clean_names()
numberOfParticiProjects <- numberOfParticiProjects %>% clean_names()
numberOfRecsProjects <- numberOfRecsProjects %>% clean_names()
numberOfPubsProjects <- numberOfPubsProjects %>% clean_names()
```

# Generate plots

## Figure 3a: Number of participants vs number of projects

Define order of factors in `numberOfParticiProjects`:

```{r}
numberOfParticiProjects$number_of_participants <- factor(numberOfParticiProjects$number_of_participants, ordered = TRUE)
```

Recode `Info not available`:

```{r}
numberOfParticiProjects <- numberOfParticiProjects %>% 
  mutate(number_of_participants = recode(number_of_participants,
                                         "Info not available" = "Unknown"))
```


Re-order `number_of_participants`:

```{r}
numberOfParticiProjects$number_of_participants <- 
  ordered(numberOfParticiProjects$number_of_participants, 
          levels = c("1 - 50", 
                     "51 - 100",
                     "101 - 500", 
                     "501 - 1000", 
                     "1001 - 2000", 
                     "2001 - 5000", 
                     "5001 - 10 000",
                     "> 10 001",
                     "Not applicable",
                     "Unknown"))
```

Produce plot:

```{r}
a <- ggplot(data = numberOfParticiProjects, 
            aes(x = number_of_participants, y = number_of_projects)) +
  geom_bar(stat="identity") +
  labs(x = "number of participants", y = "number of projects") +
  theme_classic() +
  theme(
    text = element_text(size = 30),
    axis.text = element_text(size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1)) 


png(filename = "figure3a.png",
    width = 890, height = 780)

print(a)
dev.off()
```

## Figure 3b: Number of records vs number of projects

Define order of factors in `numberOfParticiProjects`:

```{r}
numberOfRecsProjects$number_of_records <- factor(numberOfRecsProjects$number_of_records, ordered = TRUE)
```


Recode `Info not available`:

```{r}
numberOfRecsProjects <- numberOfRecsProjects %>% 
  mutate(number_of_records = recode(number_of_records,
                                    "Information not available" = "Unknown",
                                    "Not Applicable" = "Not applicable"))
```


Re-order `number_of_participants`:

```{r}
numberOfRecsProjects$number_of_records <- 
  ordered(numberOfRecsProjects$number_of_records, 
          levels = c("1 - 50", 
                     "51 - 100",
                     "101 - 500", 
                     "501 - 1000", 
                     "1001 - 2000", 
                     "2001 - 5000", 
                     "5001 - 10 000",
                     "10 001 - 50 000",
                     "50 001 - 100 000",
                     "100 001 - 500 000",
                     "500 001 - 1 000 000",
                     "> 1 000 001",
                     "Not applicable",
                     "Unknown"))
```

```{r}
b <- ggplot(data = numberOfRecsProjects, 
            aes(x = number_of_records, y = number_of_projects)) +
  geom_bar(stat="identity") +
  labs(x = "number of records", y = "number of projects") +
  theme_classic() +
  theme(
    text = element_text(size = 30),
    axis.text = element_text(size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1)) 


png(filename = "figure3b.png",
    width = 890, height = 780)

print(b)
dev.off()
```

## Figure 3c: Number of publications vs number of projects:

Define order of factors in `numberOfPubsProjects`:

```{r}
numberOfPubsProjects$number_of_publications <- factor(numberOfPubsProjects$number_of_publications, ordered = TRUE)
```

Re-order `number_of_participants`:

```{r}
numberOfPubsProjects$number_of_publications <- 
  ordered(numberOfPubsProjects$number_of_publications, 
          levels = c("0", 
                     "1 - 5",
                     "6 - 10",
                     "11 - 20", 
                     "21 - 50", 
                     "> 100",
                     "Unknown",
                     "Not Applicable"))
```

```{r}
c <- ggplot(data = numberOfPubsProjects, 
            aes(x = number_of_publications, y = number_of_projects)) +
  geom_bar(stat="identity") +
  labs(x = "number of publications", y = "number of projects") +
  theme_classic() +
  theme(
    text = element_text(size = 30),
    axis.text = element_text(size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1)) 


png(filename = "figure3c.png",
    width = 890, height = 780)

print(c)
dev.off()
```



## Figure 3d: Plot Usage Apps - Number of apps used:

Define `usage_of_app` as a factor:

```{r}
appsProjects <- appsProjects %>% mutate(usage_of_app = factor(usage_of_app, ordered = TRUE))
```

Define order of factors in `appsProjects`:

```{r}
levels(appsProjects$usage_of_app)
```
Re-order `usage_of_app`:

```{r}
appsProjects$usage_of_app <- 
  ordered(appsProjects$usage_of_app, 
          levels = c("0%", 
                     "1-5%",
                     "6-10%", 
                     "11-25%", 
                     "51-75%", 
                     "76-100%", 
                     "N/A",
                     "Unknown"))
```


```{r}
p <- ggplot(data=appsProjects, 
            aes(x=usage_of_app, 
                y=number_of_projects)) +
  geom_bar(stat="identity") + 
  labs(x = "Usage of App", y = "Number of projects") +
  theme_classic() +
  theme(
    text = element_text(size = 30),
    axis.text = element_text(size = 20),
    axis.text.x = element_text(angle = 45, hjust = 1)) 



png(filename = "figure3d.png",
    width = 890, height = 780)

print(p)
dev.off()

```

